\bf{Inequality in non-cognitive skills: \\ a cross-cohort comparison}

### 2018-05-23 codes/R/slides/cohorts_tapes.Rnw
add facineq by father occupation

\begin{frame}
\frametitle{This study}
\framesubtitle{Research questions}
\bi
\item Usually, socioemotional constructs are measured using validated scales
    \bi \item e.g. Rutter, BSAG, SDQ \ei
\pause
~\\    
\item \textbf{Q1}: Can we derive \textbf{comparable measures} of socioemotional skills across populations?
    \bi
    \item Does the same instrument measure the same constructs in the same way:
      \bi 
      \item across different populations?
      \item in the same population, at different points in time?
      \ei
    \item And what if only different instruments are available?
    \ei
\pause
~\\    
\item \textbf{Q2}: \textbf{How many latent dimensions} are underlying these instruments?
    \bi
    \item Can they be simply summarised into a single score?
    \ei
\ei
\end{frame}

@@

@@

\bi
\item Different methods for assessing number of factors give varied results \hyperlink{numfac}{\beamerbutton{Det.}}
\item We choose two factors to maximise interpretability
\item In both cohorts, EFA delivers a clear-cut split between two factors
\item Factors can be coherently interpreted as internalising and internalising skills
\ei 
\vspace{12pt}
\centering
\begin{tabular}{C{.05\textwidth}L{.3\textwidth}|C{.05\textwidth}L{.3\textwidth}}
  & \textbf{{\color{OliveGreen}Externalising}}  &    & \textbf{{\color{Orange}Internalising}}  \\
1 & Restless                & 7  & Worried                \\
2 & Squirmy/fidgety         & 8  & Afraid of new things   \\
3 & Fights/bullies children & 9  & Solitary               \\
4 & Cannot settle           & 10 & Unhappy/tearful        \\
5 & Temper tantrums         & 11 & Head/stomachaches \\
6 & Disobedient             &    &                       
\end{tabular}

@@

\bi
\item Two constructs isolated in the new scale, separately by cohort
\item But are externalising and internalising problems measured in a comparable way \emph{across cohorts}?
\vspace{12pt}\pause
\item \textbf{Measurement Invariance (MI)} analysis
  \bi
  \item Formal test of the invariance hypothesis
  \item Set up categorical Item Factor Analysis (IFA) model
  \item Pool the two cohorts together, and estimate models of different invariance level -- following \cite{Wu2016}
  \item Test relative fit of restricted models compared to initial model
  \ei 
\ei

@@

\begin{frame}
\frametitle{MI}\label{misteps}
\framesubtitle{Invariance steps \hyperlink{miapp1}{\beamerbutton{Details}}}
\bnm
\item \textbf{Configural} model
  \bi
  \item Minimally restrictive (all parameters free across groups)
  \item Normalised on latent means $\bm{\kappa}$ and variances $\bm{\Phi}$
  \ei
\item \textbf{Threshold and Loading invariance} model
  \bi
  \item Restrict $\bm{\tau}_{1}$, $\bm{\tau}_{2}$, and $\bm{\Lambda}$ to be equal across groups
  \item Allows comparison of latent factor variances
  \ei
\item \textbf{Threshold, Loading, and Intercept invariance} model
  \bi
  \item Restrict $\bm{\tau}_{1}$, $\bm{\tau}_{2}$, $\bm{\Lambda}$, and $\bm{\nu}$ to be equal across groups
  \item Allows comparison of latent factor variances \emph{and} means
  \ei
\enm
\end{frame}

@@

\begin{frame}
\frametitle{Measurement Invariance}\label{mires}
\framesubtitle{Results \hyperlink{measpars}{\beamerbutton{Measurement}}\hyperlink{lvpars}{\beamerbutton{Latent variable}}}

\bi
\item We use RMSE to compare fit between models of different invariance level
\item Restricting thresholds and loadings provides a good fit \hyperlink{fit}{\beamerbutton{Fit details}}
  \bi \item We can compare the scale (variance) of the latent factors \ei
\item Models with restricted intercepts do \emph{not} pass MI test
  \bi
  \item Unclear whether mean comparisons can be done
  \item \cite{Wu2016} vs \cite{Millsap2004}
  \ei
\ei
\begin{center}
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

@@

print(xtable(table,
                    align="L{0cm}L{.3\\textwidth}C{.1\\textwidth}C{.15\\textwidth}",
                    digits = 4
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            "Model & Num. params & RMSEA \\\\"
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "5\\textwidth",
             floating = FALSE
)

@@

\begin{frame}
\frametitle{Measurement Invariance}
\framesubtitle{Results \hyperlink{measpars}{\beamerbutton{Measurement}}\hyperlink{lvpars}{\beamerbutton{Latent variable}}}
\bi
\item \textbf{Q1}: Can we derive comparable measures of noncognitive skills across cohorts?
\item Yes, we can use the model with non-invariant intercepts to score the latent factors (externalising and internalising skills)
\item All scores are coded so that \emph{higher is better}
\ei
\end{frame}

@@

@@

\begin{frame}
\frametitle{What explains behavioural skills?}
\framesubtitle{Males}
\begin{table}
\centering
\resizebox{\textwidth}{!}{
<<REGS_M, results='asis', dependson="REGRESSIONS">>=

# select models
regs <- list(r_ext$BCS.M[["OLS"]],r_ext$MCS.M[["OLS"]],r_int$BCS.M[["OLS"]],r_int$MCS.M[["OLS"]])

@@

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = robust_se, override.pvalues = robust_p,
               single.row = TRUE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = "region|Inter",
               groups = grouplabs,
               custom.coef.names = varlabs,
               custom.model.names = c("EXT BCS", "EXT MCS", "INT BCS", "INT MCS"),
               include.rmse = FALSE,
               table = FALSE
               )

@@

\begin{frame}
\frametitle{What explains behavioural skills?}
\framesubtitle{Females}
\begin{table}
\centering
\resizebox{\textwidth}{!}{
<<REGS_F, results='asis', dependson="REGRESSIONS">>=

# select models
regs <- list(r_ext$BCS.F[["OLS"]],r_ext$MCS.F[["OLS"]],r_int$BCS.F[["OLS"]],r_int$MCS.F[["OLS"]])

@@

@
} %scalebox
\end{frame}
% 
% <<BCSEFA>>=
% @
% 
% <<BCSEFARES, dependson='BCSEFA'>>=
% @
% 
% <<MCSEFA>>=
% @
% 
% <<MCSEFARES, dependson='MCSEFA'>>=
% @
% 
% 
% \begin{frame}[noframenumbering]
% \frametitle{EFA}\label{numfac}
% \framesubtitle{Number of factors \hyperlink{efa}{\beamerreturnbutton{Back}}}
% \begin{center}
% <<EFATAB, dependson='MCSEFARES', results='asis'>>=
% # combine results
% efatab <- cbind(efatab.bcs,efatab.mcs)
% colnames(efatab) <- c("BCS All", "BCS M", "BCS F", "MCS All", "MCS M", "MCS F")
% 
% 
% print(xtable(efatab,
%                     align="L{.25\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
%                     digits = 0
%                     ),
%              size="\\small",
%              include.rownames=T,
%              include.colnames=T,
%              booktabs = T,
%              sanitize.text.function = function(x){x},
%              tabular.environment = "tabular", width = ".8\\textwidth",
%              floating = FALSE
% )
% @
% \end{center}
% \end{frame}

>>>>

\begin{frame}
\frametitle{This study}
\framesubtitle{Research questions}
\bi
\item Usually, socioemotional constructs are measured using validated scales
    \bi \item e.g. Rutter, BSAG, SDQ \ei
\pause
~\\    
\item \textbf{Q1}: Can we derive \textbf{comparable measures} of socioemotional skills across populations?
    \bi
    \item Most economic literature uses simple sum scores
    \item But does the same instrument measure the same constructs in the same way:
      \bi 
      \item across different populations?
      \item in the same population, at different points in time?
      \ei
    \item And what if only different instruments are available?
    \ei
\pause
~\\    
\item \textbf{Q2}: \textbf{How many latent dimensions} are underlying these instruments?
    \bi
    \item Can they be simply summarised into a single score?
    \ei
\ei
\end{frame}

@@

How many dimensions underlie the new 11-item scale?
\bi
\item Rutter A scale (BCS) 
  \bi 
  \item Traditionally conceptualised as bidimensional \citep{Blanden2007}
    \bi
    \item \emph{anti-social} -- or aggressive-hyperactive
    \item \emph{neurotic} -- or anxious-fearful
    \ei
  \item Sometimes the former is split into aggressive-antisocial and hyperkinetic \citep{McGee1985}
  \ei
\item Strengths and difficulties (MCS) 
  \bi
  \item Five subscales (emotional, hyperactivity, prosocial, peer, conduct) \citep{Stone2010}
  \item Evidence that lower-dimensional structures can work \citep{Goodman2010}
  \ei
\ei

@@

On our data:
\bi
\item Different methods for assessing number of factors give varied results \hyperlink{numfac}{\beamerbutton{Det.}}
\item We choose two factors to maximise interpretability
\item In both cohorts, EFA delivers a clear-cut split between two factors
\item Factors can be coherently interpreted as externalising ($\sim$ anti-social) and internalising ($\sim$ neurotic)
\ei 
\vspace{12pt}
\centering
\begin{tabular}{C{.05\textwidth}L{.3\textwidth}|C{.05\textwidth}L{.3\textwidth}}
  & \textbf{{\color{OliveGreen}Externalising}}  &    & \textbf{{\color{Orange}Internalising}}  \\
1 & Restless                & 7  & Worried                \\
2 & Squirmy/fidgety         & 8  & Afraid of new things   \\
3 & Fights/bullies children & 9  & Solitary               \\
4 & Cannot settle           & 10 & Unhappy/tearful        \\
5 & Temper tantrums         & 11 & Head/stomachaches \\
6 & Disobedient             &    &                       
\end{tabular}

@@

\bi
\item Two constructs isolated in the new scale, separately by cohort
\item But are externalising and internalising measured in a comparable way \emph{across cohorts}?
\ei
\vspace{12pt}\pause
\textbf{Measurement Invariance (MI)} analysis \citep{Wu2016a}
  \bi
  \item Formal test of the invariance hypothesis using factor analysis
  \item Aim: assess if constructs are measured in the same way in two or more groups
  \item Procedure:
    \bnm
    \item Start from the most general model -- parameters allowed differ across groups
    \item Impose restrictions on sets of parameters to be equal across groups
    \item Test relative fit of restricted models compared to initial model
    \enm
  \ei

@@

\begin{frame}
\frametitle{Measurement Invariance}\label{misteps}
\framesubtitle{Steps \hyperlink{miapp1}{\beamerbutton{Details}}}

We follow the approach in \cite{Wu2016a}
\bi
\item Define group ($c$) as cohort-gender cell
  \bi \item Invariance results do not vary if separating by gender \ei
\ei
~\\
\bnm
\item \textbf{Configural} model
  \bi
  \item Minimally restrictive (all parameters free across groups)
  \item Normalised on latent means $\bm{\kappa}$ and variances $\bm{\Phi}$
  \ei
\item \textbf{Threshold and Loading invariance} model
  \bi
  \item Restrict $\bm{\tau}_{1}$, $\bm{\tau}_{2}$, and $\bm{\Lambda}$ to be equal across groups
  \item Allows comparison of latent factor variances
  \ei
\item \textbf{Threshold, Loading, and Intercept invariance} model
  \bi
  \item Restrict $\bm{\tau}_{1}$, $\bm{\tau}_{2}$, $\bm{\Lambda}$, and $\bm{\nu}$ to be equal across groups
  \item Allows comparison of latent factor variances \emph{and} means
  \ei
\enm
\end{frame}

@@

\begin{frame}
\frametitle{Measurement Invariance}\label{mires}
\framesubtitle{Results}

\bi
\item We use RMSE to compare fit between models of different invariance level
\item Restricting thresholds and loadings provides a good fit \hyperlink{fit}{\beamerbutton{Fit details}}
  \bi \item We can compare the scale (variance) of the latent factors \ei
\item Models with restricted intercepts do \emph{not} pass MI test
  \bi \item We cannot compare the location of latent factors in different groups \ei
\ei
\begin{center}
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

@@

print(xtable(table,
                    align="L{0cm}L{.5\\textwidth}C{.1\\textwidth}C{.15\\textwidth}",
                    digits = 4
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            "Model & Num. params & RMSEA \\\\"
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@@

\begin{frame}
\frametitle{Measurement Invariance}
\framesubtitle{Factor scores}
\bi
\item Nevertheless, we can use the model with non-invariant intercepts to score the latent factors (externalising and internalising dimensions)
\item Estimated parameters: \hyperlink{measpars}{\beamerbutton{Measurement}}\hyperlink{lvpars}{\beamerbutton{Latent variable}}
\item Use Empirical Bayes approach to predict individual scores
\item All scores are coded so that \emph{higher is better}
\item \textbf{CAUTION}: location differences by cohort are not interpretable!
~\\
\pause
\item Is this an improvement on using simple sum scores?
  \bi
  \item Scores from the model exploit the covariance structure of the whole 11-item scale
  \item Every pattern of response is assigned a score
  \item Substantial increase in variability ({\color{red} could add boxplot of factor scores vs sum scores})
  \ei
\ei
\end{frame}

@@

\begin{frame}
\frametitle{Inequality}\label{ineq_fsc}
\framesubtitle{Father occupation}
\vspace{-.2cm}
\begin{figure}
<<FACINEQ_FSC, dependson="FA_SCORES_BIND", dev='tikz', fig.height=7.5, fig.width=10, out.width='.9\\linewidth'>>=
# plot
fi <- fi_fsc # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{What explains socioemotional skills?}
\framesubtitle{Externalising}
\begin{table}
\centering
\resizebox{\textwidth}{!}{
<<REGS_M, results='asis', dependson="REGRESSIONS">>=

# select models
regs <- list(r_ext$BCS.M[["OLS"]],r_ext$MCS.M[["OLS"]],r_ext$BCS.F[["OLS"]],r_ext$MCS.F[["OLS"]])

@@

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = robust_se, override.pvalues = robust_p,
               single.row = TRUE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = "region|Inter",
               groups = grouplabs,
               custom.coef.names = varlabs,
               custom.model.names = c("Males BCS", "Males MCS", "Females BCS", "Females MCS"),
               include.rmse = FALSE,
               table = FALSE
               )

@@

\begin{frame}
\frametitle{What explains socioemotional skills?}
\framesubtitle{Internalising}
\begin{table}
\centering
\resizebox{\textwidth}{!}{
<<REGS_F, results='asis', dependson="REGRESSIONS">>=

# select models
regs <- list(r_int$BCS.M[["OLS"]],r_int$MCS.M[["OLS"]],r_int$BCS.F[["OLS"]],r_int$MCS.F[["OLS"]])

@@

<<BCSEFA>>=
@

<<BCSEFARES, dependson='BCSEFA'>>=
@

<<MCSEFA>>=
@

<<MCSEFARES, dependson='MCSEFA'>>=
@

\begin{frame}[noframenumbering]
\frametitle{EFA}\label{numfac}
\framesubtitle{Number of factors \hyperlink{efa}{\beamerreturnbutton{Back}}}
\begin{center}
<<EFATAB, dependson='MCSEFARES', results='asis'>>=
# combine results
efatab <- cbind(efatab.bcs,efatab.mcs)
colnames(efatab) <- c("BCS All", "BCS M", "BCS F", "MCS All", "MCS M", "MCS F")

print(xtable(efatab,
                    align="L{.25\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
                    digits = 0
                    ),
             size="\\small",
             include.rownames=T,
             include.colnames=T,
             booktabs = T,
             sanitize.text.function = function(x){x},
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)
@
\end{center}
\end{frame}
